---
output: github_document
---

<!-- Render README.Rmd. Then open README.md and use VS Code (Markdown All in One) to create/update TOC. -->

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE
)
```

# r-windows-paths

Essential R command-line tool paths on Windows.
This information can help setting up the `PATH` environment variable
for seamless command-line access.

## Contents

## Example versions

Some of the paths below are specific to the installed software version.
We use these software version examples:

```{r}
r_version <- "4.4.1"
rtools_version <- "44"
rstudio_version_major <- "2024.04.2"
rstudio_version_minor <- "764"
```

- R `r r_version`
- Rtools`r rtools_version`
- RStudio Desktop `r sprintf("%s+%s", rstudio_version_major, rstudio_version_minor)`

## Key paths

### R

`R.exe` and `Rscript.exe`.

R release version `r r_version`:

```powershell
C:\Program Files\R\R-`r r_version`\bin
```

R-devel:

```powershell
C:\Program Files\R\R-devel\bin
```

### Rtools

C/C++/Fortran toolchain for building R and R packages from source.

Contains `gcc`, `g++`, `gfortran`:

```powershell
C:\rtools`r rtools_version`\x86_64-w64-mingw32.static.posix\bin
```

Contains Texinfo (`texi2dvi`, `texi2pdf`) required by `R CMD check`:

```powershell
C:\rtools`r rtools_version`\usr\bin
```

### Pandoc

pandoc installed via `choco install pandoc`:

```powershell
C:\Users\%USERNAME%\AppData\Local\Pandoc
```

pandoc distributed with RStudio Desktop:

```powershell
C:\Users\%USERNAME%\RStudio-`r sprintf("%s-%s", rstudio_version_major, rstudio_version_minor)`\resources\app\bin\quarto\bin\tools
```

pandoc distributed with Quarto standalone installer:

```powershell
C:\Users\%USERNAME%\AppData\Local\Programs\Quarto\bin\tools
```

### TinyTeX

TeX distribution binaries for LaTeX support.
Contains `pdflatex` required by `R CMD check`.

```powershell
C:\Users\%USERNAME%\AppData\Roaming\TinyTeX\bin\windows
```

### Quarto

Quarto distributed with RStudio Desktop:

```powershell
C:\Users\%USERNAME%\RStudio-`r sprintf("%s-%s", rstudio_version_major, rstudio_version_minor)`\resources\app\bin\quarto\bin
```

Quarto distributed via Quarto standalone installer:

```powershell
C:\Users\%USERNAME%\AppData\Local\Programs\Quarto\bin
```

## How to use

<!-- TODO: Add screenshots -->

1. Locate the paths on your system.
1. Add the relevant paths to your system's `PATH` environment variable.
1. Verify the setup by running the commands in command prompt.
